### 2021-05-21 R0-Density-Reanalysis/R0-Density-Reanalysis.tex

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{McMaster University}
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{The importance of reproducibility in COVID-19 research:
the case of population density and the spread of the
pandemic} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Antonio Paez\textsuperscript{\getAff{McMaster
University}}\textsuperscript{*}\\
\bigskip
\textbf{\getAff{McMaster University}}School of Earth, Environment and
Society, 1280 Main St West, Hamilton, Ontario L8S 4K1 Canada\\
\bigskip
* Corresponding author: paezha@mcmaster.ca\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
The emergence of the novel SARS-CoV-2 coronavirus and the global
COVID-19 pandemic has led to explosive growth in scientific research.
Given the high stakes of the situation, it is essential that scientific
activites, on which good policy depends, are as transparent and
reproducible as possible. Reproducibility is key for the efficient
operation of the self-correction mechanisms of science, which work to
weed out errors and refine our understanding of social and physical
phenomena. In this paper, the importance of reproducibility is
illustrated for the case of the association between population density
and the the spread of SARS-CoV-2. Transparency and openness means that
the same problem can, with relatively modest efforts, be examined by
independent researchers who can verify findings, and bring to bear
different perspectives, approaches, and methods---sometimes with
consequantial changes in the conclusions, as the empirical example in
this paper shows.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The emergence of the novel SARS-CoV-2 coronavirus in 2019, and the
global pandemic that followed in its wake, led to an explosive growth of
research around the globe. According to Fraser et al. {[}1{]}, over
125,000 COVID-19-related papers were released in the first ten months
from the first confirmed case of the disease. Of these, more than 30,000
were shared in pre-print servers, the use of which also exploded in the
past year {[}2--4{]}.

Given the heavy human and economic cost of the pandemic, there has been
a natural tension in the scientific community between the need to
publish research results quickly and the imperative to maintain
consistently high quality standards in scientific reporting; indeed, a
call for maintaining the standards in published research has even called
this deluge of publications a ``carnage of substandard research''
{[}5{]}. Part of the challenge of maintaining quality standards in
published research is that, despite an abundance of recommendations and
guidelines {[}6--9{]}, in practice reproducibility has remained a lofty
and somewhat aspirational goal {[}10,11{]}. As reported in the
literature, only a woefully small proportion of published research was
actually reproducible before the pandemic {[}12,13{]}, and the situation
does not appear to have changed substantially since {[}14,15{]}.

The push for open data and software, along with more strenuous efforts
towards open, reproducible research, is simply a continuation of
long-standing scientific practices of independent verification. Despite
the (at times disproportionate) attention that high profile scandals in
science tend to elicit in the media, science as a collective endeavor is
remarkable for being a self-correcting enterprise, one with built-in
mechanisms and incentives to weed out erroneous ideas. Over the long
term, facts tend to prevail in science. At stake is the shorter-term
impacts that research may have in other spheres of economic and social
life. The case of economists Reinhart and Rogoff comes to mind: by the
time the inaccuracies and errors in their research were uncovered {[}see
16{]}, their claims about debt and economic growth had already been
seized by policy-makers on both sides of the Atlantic to justify
austerity policies in the aftermath of the Great Recession of
2007-2009\footnote{Nobel Prize in Economics Paul Krugman noted that
  ``Reinhart--Rogoff may have had more immediate influence on public
  debate than any previous paper in the history of economics''
  \url{https://www.nybooks.com/articles/2013/06/06/how-case-austerity-has-crumbled/?pagination=false}}.
As later research has demonstrated, those policies cast a long shadow,
and their sequels continued to be felt for years {[}17{]}.

In the context of COVID-19, a topic that has grabbed the imagination of
numerous thinkers has been the prospect of life in cities after the
pandemic {[}18{]}. The fact that the worst of the pandemic was initially
felt in dense population centers such as Wuhan, Milan, Madrid, and New
York, brought a torrent of research into the associations between
density and the spread of the pandemic. Some important questions hang on
the results of these research efforts. For example, are lower density
regions safer from the pandemic? Are de-densification policies
warranted, even if just in the short term? And in the longer term, will
the risks of life in high density regions presage a flight from cities?
Over the past year, numerous papers have sought to throw light into the
underlying issue of density and the pandemic; nonetheless the results,
as will be detailed next, remain mixed. Further, to complicate matters,
precious few of these studies appear to be sufficiently open to support
independent verification.

The objective of this paper is to illustrate the importance of
reproducibility in research, particularly in the context of the flood of
COVID-19 papers. To this end, a recent study by Sy et al. {[}19{]} is
chosen as an example of reproducible research. The objective is not to
malign the analysis of these researchers, but rather to demonstrate the
value of openness to allow for independent verification and further
analysis. Open data and open code mean that an independent researcher
can, with only modest efforts, not only verify the findings reported,
but also examine the same data from a perspective which may not have
been available to the original researchers due to differences in
disciplinary perspectives, methodological traditions, and/or training,
among other possible factors. The example, which shows consequential
changes in the conclusions reached by different analyses, should serve
as a call to researchers to redouble their efforts to increase
transparency and reproducibility in research. This paper, in addition,
aims to show how data can be packaged in well-documented, shareable
units, and code can be embedded into self-contained documents suitable
for review and independent verification. The source for this paper is an
\href{http://rmarkdown.rstudio.com}{R Markdown} document which, along
with the data package, is available in a public repository\footnote{\url{https://github.com/paezha/Reproductive-Rate-and-Density-US-Reanalyzed}}.

\hypertarget{background-the-intuitive-relationship-between-density-and-spread-of-contagious-diseases}{%
\section{Background: the intuitive relationship between density and
spread of contagious
diseases}\label{background-the-intuitive-relationship-between-density-and-spread-of-contagious-diseases}}

The concern with population density and the spread of the virus during
the COVID-19 pandemic was fueled, at least in part, by dramatic scenes
seen in real-time around the world from large urban centers such as
Wuhan, Milan, Madrid, and New York. In theory, there are good reasons to
believe that higher density may have a positive association with the
transmission of a contagious virus. It has long been known that the
potential for inter-personal contact is greater in regions with higher
density {[}see for example the research on urban fields and
time-geography 20,21,22{]}. Mathematically, models of exposure and
contagion indicate that higher densities can catalyze the transmission
of contagious diseases {[}23,24{]}. The idea is intuitive and likely at
the root of messages, by some figures in positions of authority, that
regions with sparse population densities faced lower risks from the
pandemic\footnote{Governor Kristi Noem of South Dakota, for example,
  claimed that sparse population density allowed her state to face the
  pandemic down without the need for strict policy interventions
  \url{https://www.inforum.com/lifestyle/health/5025620-South-Dakota-is-not-New-York-City-Noem-defends-lack-of-statewide-COVID-19-restrictions}}.

As Rocklöv and Sjödin {[}23{]} note, however, mathematical models of
contagion are valid at small-to-medium spatial scales (and presumably,
small temporal scales too, such as time spent in restaurants, concert
halls, cruises), and the results do not necessarily transfer to larger
spatial units and different time scales. There are solid reasons for
this: while in a restaurant, one can hardly avoid being in proximity to
other customers-however, a person can choose to (or be forced to as a
matter of policy) not go to a restaurant in the first place.
Nonetheless, the idea that high density correlates with high
transmission is so seemingly reasonable that it is often taken for
granted even at larger scales {[}e.g., 25,26{]}. At larger scales,
however, there exists the possibility of behavioral adaptations, which
are difficult to capture in the mechanistic framework of differential
equations {[}or can be missing in agent-based models, e.g., 27{]}; these
adaptations, in fact, can be a key aspect of disease transmission.

A plausible behavioral adaptation during a pandemic, especially one
broadcast as widely and intensely as COVID-19, is risk compensation.
Risk compensation is a process whereby people adjust their behavior in
response to their \emph{perception} of risk {[}28--30{]}. In the case of
COVID-19, Chauhan et al. {[}31{]} have found that perception of risks in
the US varies between rural, suburban, and urban residents, with rural
residents in general expressing less concern about the virus. It is
possible that people who listened to the message of leaders saying that
they were safe because of low density may not have taken adequate
precautions against the virus. People in dense places who could more
directly observe the impact of the pandemic may have become overly
cautious. Both Paez et al. {[}32{]} and Hamidi et al. {[}33{]} posit
this mechanism (i.e., greater compliance with social distancing in
denser regions) to explain the results of their analyses. The evidence
available does indeed show that there were important changes in behavior
with respect to mobility during the pandemic {[}34--36{]}; furthermore,
shelter in place orders may have had greater buy-in from the public in
higher density regions {[}37{]}, and the associated behavior may have
persisted beyond the duration of official social-distancing policies
{[}38{]}. In addition, there is evidence that changes in mobility
correlated with the trajectory of the pandemic {[}39,40{]}. Given the
potential for behavioral adaptation, the question of density becomes
more nuanced: it is not just a matter of proximity, but also of human
behavior, which is better studied using population-level data and
models.

\hypertarget{background-but-what-does-the-literature-say}{%
\section{Background: but what does the literature
say?}\label{background-but-what-does-the-literature-say}}

When it comes to population density and the spread of COVID-19, the
international literature to date remains inconclusive.

On the one hand, there are studies that report positive associations
between population density and various COVID-19-related outcomes. Bhadra
{[}41{]}, for example, reported a moderate positive correlation between
the spread of COVID-19 and population density at the district level in
India, however their analysis was bivariate and did not control for
other variables, such as income. Similarly, Kadi and Khelfaoui {[}42{]}
found a positive and significant correlation between number of cases and
population density in cities in Algeria in a series of simple regression
models (i.e., without other controls). A question in these relatively
simple analyses is whether density is not a proxy for other factors.
Other studies have included controls, such as Pequeno et al. {[}43{]}, a
team that reported a positive association between density and cumulative
counts of confirmed COVID-19 cases in state capitals in Brazil after
controlling for covariates, including income, transport connectivity,
and economic status. In a similar vein, Fielding-Miller et al. {[}44{]}
reported a positive relationship between the absolute number of COVID-19
deaths and population density (rate) in rural counties in the US. Roy
and Ghosh {[}45{]} used a battery of machine learning techniques to find
discriminatory factors, and a positive and significant association
between COVID-19 infection and death rates in US states. Wong and Li
{[}46{]} also found a positive and significant association between
population density and number of confirmed COVID-19 cases in US
counties, using both univariate and multivariate regressions with
spatial effects. More recently, Sy et al. {[}19{]} reported that the
basic reproductive number of COVID-19 in US counties tended to increase
with population density, but at a decreasing rate at higher densities.

On the flip side, a number of studies report non-significant or negative
associations between population density and COVID-19 outcomes. This
includes the research of Sun et al. {[}47{]} who did not find evidence
of significant correlation between population density and confirmed
number of cases per day \emph{in conditions of lockdown} in China. This
finding echoes the results of Paez et al. {[}32{]}, who in their study
of provinces in Spain reported non-significant associations between
population density and infection rates in the early days of the first
wave of COVID-19, and negative significant associations in the later
part of the first lockdown. Similarly, {[}48{]} found zero or negative
associations between population density and infection numbers/deaths by
country. Fielding-Miller et al. {[}44{]} contrast their finding about
rural counties with a negative relationship between COVID-19 deaths and
population density in urban counties in the US. For their part, in their
investigation of doubling time, White and Hébert-Dufresne {[}49{]}
identified a negative and significant correlation between population
density and doubling time in US states. Likewise, {[}50{]} fond a small
negative (and significant) association between population density and
COVID-19 morbidity in districts in Tehran. Finally, two of the most
complete studies in the US {[}by Hamidi et al.; {[}51{]}; {[}33{]}{]}
used an extensive set of controls to find negative and significant
correlations between density and COVID-19 cases and fatalities at the
level of counties in the US.

As can be seen, these studies are implemented at different scales in
different regions of the world. They also use a range of techniques,
from correlation analysis, to multivariate regression, spatial
regressions, and machine learning techniques. This is natural and to be
expected: individual researchers have only limited time and expertise.
This is why reproducibility is important. To pick an example (which will
be further elaborated in later sections of this paper), the study of Sy
et al.~{[}{[}19{]}; hereafter SWN{]} would immediately grab the
attention of a researcher with a somewhat different toolbox.

\hypertarget{reproducibility-of-research}{%
\section{Reproducibility of
research}\label{reproducibility-of-research}}

SWN investigated the basic reproductive number of COVID-19 in US
counties, and its association with population density, median household
income, and prevalence of private mobility. For their multivariate
analysis, SWN used mixed linear models. This is a reasonable modelling
choice: \(R_0\) is an interval-ratio variable that is suitably modeled
using linear regression; further, as SWN note there is a likelihood that
the process in not independent ``among counties within each state,
potentially due to variable resource allocation and differing health
systems across states'' (p.~3). A mixed linear model accounts for this
by introducing random components (in the case of SWN, random intercepts
at the state level). SWN estimated various models with different
combinations of variables, including median household income and
prevalence of travel by private transportation. These are sensible
controls, given potential variations in behavior: people in more
affluent counties may have greater opportunities to work from home, and
use of private transportation reduces contact with strangers. Moreover,
they also conducted various sensitivity analyses. After these efforts,
SWN conclude that there is a positive association between the basic
reproductive number and population density at the level of counties in
the US.

One salient aspect of the analysis in SWN is that the basic reproductive
number can only be calculated reliably with a minimum number of cases,
and a large number of counties did not meet such threshold. As
researchers do, SWN made modelling decisions, in this case basing their
analysis only on counties with valid observations. A modeler with
expertise in different methods would likely ask some of the following
questions on reading SWN's paper: how were missing counties treated?
What are the implications of the spatial sampling framework used in the
analysis? Is it possible to spatially interpolate the missing
observations? And, was there evidence of spatial autocorrelation in the
residuals of the models? These questions are relevant and their
implications important. Fortunately, SWN are an example of a reasonably
open, reproducible research product: their paper is accompanied by (most
of) the data and (most of) the code used in the analysis. This means
that an independent expert can, with only a moderate investment of time
and effort, reproduce the results in the paper, as well as ask
additional questions.

Alas, reproducibility is not necessarily the norm in the relevant
literature.

There are various reasons why a project can fail to be reproducible. In
some cases, there might be legitimate reasons to withhold the data,
perhaps due to confidentiality and privacy reasons {[}e.g., 52{]}. But
in many other cases the data are publicly available, which in fact has
commonly been the case with population-level COVID-19 information.
Typically the provenance of the data is documented, but in numerous
studies the data themselves are not shared {[}25,33,41,44,51,53--56{]}.
As any researcher can attest, whether a graduate student or a seasoned
scientist, collecting, organizing, and preparing data for a project can
take a substantial amount of time. Pointing to the sources of data, even
when these sources are public, is a small step towards
reproducibility-but only a very small one. Faced with the prospect of
having to recreate a data set from raw sources is probably sufficient to
dissuade all but the most dedicated (or stubborn) researcher from
independent verification. This is true even if part of the data are
shared {[}e.g., 46{]}. In other cases, data are shared, but the
processes followed in the preparation of the data are not fully
documented {[}48,57{]}. These processes matter, as shown by the errors
in the spreadsheets of Reinhart and Rogoff {[}16{]} and the data of
biologist Jonathan Pruitt that led to an ``avalanche'' of paper
retractions\footnote{\url{https://doi.org/10.1038/d41586-020-00287-y}}.
Another situation is when papers share well-documented data, but fail to
provide the code used in the analysis {[}43,58,59{]}. Making code
available only ``on demand'' {[}e.g., 60{]} is an unnecessary barrier
when most journals offer the facility to share supplemental materials
online. Then there are those papers that more closely comply with
reproducibility standards, and share well-documented processes and data,
as well as the code used in any analyses reported {[}19,32,37,49,61{]}.

In the following sections, the analysis of RWN is reproduced, some
relevant questions from the perspective of a spatial modeler are asked,
and the data are reanalyzed.

\hypertarget{reproducing-swn}{%
\section{Reproducing SWN}\label{reproducing-swn}}

SWN examined the association between the basic reproductive number of
COVID-19 and population density. The basic reproductive number \(R_0\)
is a summary measure of contact rates, probability of transmission of a
pathogen, and duration of infectiousness. In rough terms, \(R_0\)
measures how many new infections each infections begets. Infectious
disease outbreaks generally tend to die out when \(R_0<1\), and to grow
when \(R_0>1\). Reliable calculation of \(R_0\) requires a minimum
number of cases to be able to assume that there is community
transmission of the pathogen. Accordingly, SWN based their analysis only
on counties that had at least 25 cases or more at the end of the
exponential growth phase (see Fig. \ref{fig:R0-map}). Their final sample
included 1,151 counties in the US, including in Alaska, Hawaii, Puerto
Rico, and island territories.

\begin{figure}
\includegraphics[width=1\linewidth]{R0-Density-Reanalysis_files/figure-latex/R0-map-1} \caption{\label{fig:R0-map}Basic reproductive rate in US counties (Alaska, Hawaii, Puerto Rico, and territories not shown).}\label{fig:R0-map}
\end{figure}

Table \ref{tab:swn-results} reproduces the first three models of SWN
(the fourth model did not have any significant variables; see Table 1 in
SWN). It is possible to verify that the results match, with only the
minor (and irrelevant) exception of the magnitude of the coefficient for
travel by private transportation, which is due to a difference in the
input (here the variable is one percent units, whereas in SWN it was ten
percent units). The mixed linear model gives random intercepts (i.e.,
the intercept is a random variable), and the standard deviation is
reported in the fourth row of Table \ref{tab:swn-results}. It is useful
to map the random intercepts: as seen in Figure
\ref{fig:random-terms-map}, other things being equal, counties in Texas
tend to have somewhat lower values of \(R_0\) (i.e., a negative random
intercept), whereas counties in South Dakota tend to have higher values
of \(R_0\). The key of the analysis, after extensive sensitivity
analysis, is a robust finding that population density has a positive
association with the basic reproductive number. But does it?

\begin{table}

\caption{\label{tab:tabulate-swn-results}\label{tab:swn-results}Reproducing SWN: Models 1-3}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllllrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Model 1} & \multicolumn{2}{c}{Model 2} & \multicolumn{2}{c}{Model 3} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7}
Variable & beta & 95\% CI & beta & 95\% CI & beta & 95\% CI\\
\midrule
Intercept & 2.274 & [2.167, 2.381] & 3.347 & [2.676, 4.018] & 3.386 & [2.614, 4.157]\\
Log of population density & 0.162 & [0.133, 0.191] & 0.145 & [0.115, 0.176] & 0.147 & [0.113, 0.18]\\
Percent of private transportation &  &  & -0.013 & [-0.02, -0.005] & -0.013 & [-0.021, -0.005]\\
Median household income (\$10,000) &  &  &  &  & -0.003 & [-0.033, 0.026]\\
Standard deviation (Intercept) & 0.166 & [0.108, 0.254] & 0.136 & [0.081, 0.229] & 0.137 & [0.081, 0.232]\\
\addlinespace
Within-group standard error & 0.665 & [0.638, 0.693] & 0.665 & [0.638, 0.693] & 0.665 & [0.638, 0.694]\\
\bottomrule
\end{tabular}}
\end{table}

\begin{figure}
\includegraphics[width=1\linewidth]{R0-Density-Reanalysis_files/figure-latex/random-terms-map-1} \caption{\label{fig:random-terms-map}Random intercepts of Model 3 (Alaska, Hawaii, Puerto Rico, and territories not shown).}\label{fig:random-terms-map}
\end{figure}

\hypertarget{expanding-on-swn}{%
\section{Expanding on SWN}\label{expanding-on-swn}}

The preceding section shows that thanks to the availability of code and
data, it is possible to verify the results reported by SWN. As noted
earlier, though, an independent researcher might have wondered about the
implications of the spatial sampling procedure used by SWN. The decision
to use a sample of counties with reliable basic reproductive numbers,
although apparently sensible, results in a non-random spatial sampling
scheme. Turning our attention back to Figure \ref{fig:R0-map}, we form
the impression that many counties without reliable values of \(R_0\) are
in more rural, less dense parts of the United States. This impression is
reinforced when we overlay the boundaries of urban areas with population
greater than 50,000 on the counties with valid values of \(R_0\) (see
Figure \ref{fig:urban-areas-map}). The fact that \(R_0\) could not be
accurately computed in many counties without large urban areas does not
mean that there was no transmission of the virus: it simply means that
we do not know with precision whether that was the case. The low number
of cases may be related to low population and/or low population density.
This is intriguing, to say the least: by excluding cases based on the
ability to calculate \(R_0\) we are potentially \emph{censoring} the
sample in a non-random way.

\begin{figure}
\includegraphics[width=1\linewidth]{R0-Density-Reanalysis_files/figure-latex/urban-areas-map-1} \caption{\label{fig:urban-areas-map}Urban areas with population > 50,000 (Alaska, Hawaii, Puerto Rico, and territories not shown).}\label{fig:urban-areas-map}
\end{figure}

A problematic issue with non-random sample selection is that parameter
estimates can become unreliable, and numerous techniques have been
developed over time to address this. A model useful for sample selection
problems is Heckman's selection model {[}see 62{]}. The selection model
is in fact a system of two equations, as follows:

\[
\begin{array}{c}
y_i^{S*} = \beta^{S\prime}x_i^S+\epsilon_i^S\\
y_i^{O*} = \beta^{O\prime}x_i^O+\epsilon_i^O
\end{array}
\] \noindent where \(y_i^{S*}\) is a latent variable for the sample
selection process, and \(y_i^{O*}\) is the latent outcome. Vectors
\(x_i^S\) and \(x_i^O\) are explanatory variables (with the possibility
that \(x_i^S = x_i^S\)). Both equations include random terms (i.e.,
\(\epsilon_i^S\) and \(\epsilon_i^O\)) The first equation is designed to
model the \emph{probability} of sampling, and the second equation the
outcome of interest (say \(R_0\)). The random terms are jointly
distributed and correlated with parameter \(\rho\).

What the analyst observes is the following: \[
y_i^S =
\begin{cases}
0 & \text{if } y_i^{S*} < 0\\
1 & \text{otherwise}
\end{cases}
\] \noindent and: \[
y_i^O =
\begin{cases}
0 & \text{if } y_i^{S} = 0\\
y_i^{O*} & \text{otherwise}
\end{cases}
\]

In other words, the outcome of interest is observed \emph{only} for
certain cases (\(y_i^S=1\), i.e., for sampled observations). The
probability of sampling depends on \(x_i^S\). For the cases observed,
the outcome \(y_i^O\) depends on \(x_i^O\).

A sample selection model is estimated using the same selection of
variables as SWN Model 3. This is Sample Selection Model 1 in Table
\ref{tab:selection-results}. The first thing to notice about this model
is that the sample selection process and the outcome are not independent
(\(\rho\ne0\) with 5\% of confidence). The selection equation indicates
that the probability of a county to be in the sample increases with
population density (but at a decreasing rate due to the
log-transformation), when travel by private modes is more prevalent, and
as median household income in the county is higher. This is in line with
the impression left by Figure \ref{fig:urban-areas-map} that counties
with reliable values of \(R_0\) tended to be those with larger urban
centers. Once that the selection probabilities are accounted for in the
model, several things happen with the outcomes model. First, the
coefficient for population density is still positive, but the magnitude
changes: in effect, it appears that the effect of density is more
pronounced than what SWN Model 3 indicated. The coefficient for percent
of private transportation changes signs. And the coefficient for median
household income is now significant.

The second model in Table \ref{tab:selection-results} (Selection Model
2) changes the way the variables are entered into the model. The
log-transformation of density in SWN and Selection Model 1 assumes that
the association between density and \(R_0\) is monotonically increasing
(if the sign of the coefficient is positive) or decreasing (if the sign
of the coefficient is negative). There are some indications that the
relationship may actually not be monotonical. For example, Paez et al.
{[}32{]} found a positive (if non-significant) relationship between
density and incidence of COVID-19 in the provinces of Spain at the
beginning of the pandemic. This changed to a negative (and significant)
relationship during the lockdown. In the case of the US, Fielding-Miller
et al. {[}44{]} found that the association between COVID-19 deaths and
population density was positive in rural counties, but negative in urban
counties. A variable transformation that allows for non-monotonic
changes in the relationship is the square of the density.

As seen in the table, Selection Model 2 replaces the log-transformation
of population density with a quadratic expansion. The results of this
analysis indicate that with this variable transformation, the selection
and outcome processes are still not independent (\(\rho\ne0\) with 5\%
of confidence). But a few interesting things emerge. When we examine the
outcomes model, we see that the quadratic expansion has a positive
coefficient for the first order term, but a negative coefficient for the
second order term. This indicates that \(R_0\) initially tends to
increase with higher density, but only up to a point, after which the
negative second term (which grows more rapidly due to the square),
becomes increasingly dominant. Secondly, the sign of the coefficient for
travel by private transportation becomes negative again. This, of
course, makes more sense than the positive sign of Selection Model 1: if
people tend to travel in private transportation, the potential for
contact should be lower instead of higher. And finally median household
income is no longer significant.

\begin{table}

\caption{\label{tab:tabulate-sample-selection-results}\label{tab:selection-results}Estimation results of sample selection models}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Selection Model 1} & \multicolumn{2}{c}{Selection Model 2} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
Variable & $\beta$ & 95\% CI & $\beta$ & 95\% CI\\
\midrule
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Sample Selection Model}}\\
\hspace{1em}Intercept & -2.237 & [-3.109, -1.365] & -7.339 & [-8.381, -6.297]\\
\hspace{1em}Log of population density & 0.385 & [0.352, 0.418] &  & \\
\hspace{1em}Density (1,000 per sq.km) &  &  & 2.484 & [2.13, 2.838]\\
\hspace{1em}Density squared &  &  & -0.387 & [-0.473, -0.3]\\
\hspace{1em}Percent of private transportation & 0.025 & [0.016, 0.034] & 0.057 & [0.046, 0.067]\\
\hspace{1em}Median household income (10,000) & 0.202 & [0.168, 0.235] & 0.32 & [0.283, 0.357]\\
\addlinespace[0.3em]
\multicolumn{5}{l}{\textbf{Outcome Model}}\\
\hspace{1em}Intercept & 0.605 & [-0.257, 1.466] & 2.784 & [1.652, 3.915]\\
\hspace{1em}Log of population density & 0.39 & [0.354, 0.426] &  & \\
\hspace{1em}Density (1,000 per sq.km) &  &  & 0.758 & [0.509, 1.008]\\
\hspace{1em}Density squared &  &  & -0.132 & [-0.187, -0.077]\\
\hspace{1em}Percent of private transportation & 0.01 & [0.001, 0.018] & -0.011 & [-0.021, -0.001]\\
\hspace{1em}Median household income (\$10,000) & 0.126 & [0.094, 0.159] & 0.002 & [-0.033, 0.037]\\
$\sigma$ & 0.954 & [0.904, 1.003] & 0.684 & [0.652, 0.716]\\
$\rho$ & 0.971 & [0.961, 0.98] & -0.199 & [-0.377, -0.022]\\
\bottomrule
\end{tabular}}
\end{table}

How relevant is the difference between these different model
specifications? Figure \ref{fig:comparison-results} shows the
relationship between density and \(R_0\) implied by SWN Model 1 and
Selection Model 2. The left panel of the figure shows the non-linear but
monotonic relationship implied by SWN Model 1. The conclusion is that at
higher densities, \(R_0\) is \emph{always} higher. The right panel, in
contrast, shows that, according to Selection Model 2, \(R_0\) is zero
when density is zero (as expected), and then it tends to increase at
higher densities. This continues until a density of approximately 2.9
(1,000 people per sq.km). At higher densities than that \(R_0\) begins
to decline, and the relationship becomes negative at densities higher
than approximately 5.7 (1,000 people per sq.km).

Thus, other things being equal, the effect of density in a county like
Charlottesville in Virginia (density \textasciitilde1,639 people per
sq.km) is roughly the same as that in a county like Philadelphia
(density \textasciitilde4,127 people per sq.km). In contrast, the effect
of density on \(R_0\) in a county like Arlington in Virginia (density
\textasciitilde3,093 people per sq.km) is \emph{stronger} than either of
the previous two examples. Lastly, the density of counties like San
Francisco in California, or Queens and Bronx in NY, which are among the
densest in the US, contributes even less to \(R_0\) than even the most
rural counties in the country.

It is worth at this point to recall Cressie's dictum about modelling:
``{[}w{]}hat is one person's mean structure could be another person's
correlation structure'' {[}63{]}. There are almost always multiple ways
to approach a modelling situation. In the present case, we would argue
that spatial sampling is an important aspect of the modeling process,
but one that perhaps required different technical skills than those
available to SWN. There is nothing wrong with that. What matters is
that, by adopting relatively high reproducibility standards, these
researchers made a valuable and honest contribution to the collective
enterprise of seeking knowledge. Their effort, and subsequent efforts to
validate and expand on their work, can potentially contribute to provide
clarity to ongoing conversations about the relevance of density and the
spread of COVID-19.

In particular, it is noteworthy that a sample selection model with a
different variable transformation does not lend support to the thesis
that higher density is \emph{always} associated with a greater risk of
spread of the virus {[}as put by Wong and Li, ``\,`Density is destiny'
is probably an overstatement''; {[}46{]}{]}. At the same time, this also
stands in contrast to the findings of Hamidi et al., who found that
higher density was either not significantly associated with the rate of
the virus in a cross-sectional study {[}33{]}, or was negatively
associated with in a longitudinal setting {[}{[}51{]}. In this sense,
the conclusion that density does not aggravate the pandemic may have
been somewhat premature; instead, reanalysis of the data of SWN suggests
that Fielding-Miller et al. {[}44{]} might be onto something with
respect to the difference between rural and urban counties. More
generally, in population-level studies, density is indicative of
proximity, no doubt about that, but also for adaptive behavior. And it
is possible that the determining factor during COVID-19, at least in the
US, has been variations in perceptions of the risks associated with
contagion {[}31{]}, and subsequent compensations in behavior in more and
less dense regions.

\begin{figure}
\includegraphics[width=1\linewidth]{R0-Density-Reanalysis_files/figure-latex/comparison-results-1} \caption{\label{fig:comparison-results}Effect of density according to SWN Model 3 and Sample Selection Model 2.}\label{fig:comparison-results}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

The tension between the need to publish research potentially useful in
dealing with a global pandemic, and a ``carnage of substandard
research'' {[}5{]}, highlights the importance of efforts to maintain the
quality of scientific outputs during COVID-19. An important part of
quality control is the ability of independent researchers to verify and
examine the results of materials published in the literature. As
previous research illustrates, reproducibility in scientific research
remains an important but elusive goal {[}e.g., 12,13--15{]}. This idea
is reinforced by the review conducted for this paper in the context of
research about population density and the spread of COVID-19.

Taking one recent example from the literature {[}Sy et al., {[}19{]};
SWN{]}, the present paper illustrates the importance of good
reproducibility practices. Sharing data and code can catalyze research,
by allowing independent verification of findings, as well as additional
research. After verifying the results of SWN, experiments with sample
selection models and variations in the definition of model inputs, lead
to an important reappraisal of the conclusion that high density is
associated with greater spread of the virus. Instead, the possibility of
a non-monotonical relationship between population density and contagion
is raised.

In the spirit of openness, this paper is prepared as an R Markdown
document, an a companion data package is provided. The data package
contains the relevant documentation of the data, and all data
pre-processing is fully documented. Hopefully this, and similar
reproducible papers, will continue to encourage others to adopt
reproducible standards in their research.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\hypertarget{ref-Fraser2021evolving}{}%
\CSLLeftMargin{1. }
\CSLRightInline{Fraser N, Brierley L, Dey G, Polka JK, Pálfy M, Nanni F,
et al. The evolving role of preprints in the dissemination of COVID-19
research and their impact on the science communication landscape. PLOS
Biology. 2021;19: e3000959.
doi:\href{https://doi.org/10.1371/journal.pbio.3000959}{10.1371/journal.pbio.3000959}}

\leavevmode\hypertarget{ref-Kwon2021swamped}{}%
\CSLLeftMargin{2. }
\CSLRightInline{Kwon D. How swamped preprint servers are blocking bad
coronavirus research. Nature. 2020;581: 130--132. }

\leavevmode\hypertarget{ref-Vlasschaert2020proliferation}{}%
\CSLLeftMargin{3. }
\CSLRightInline{Vlasschaert C, Topf JM, Hiremath S. Proliferation of
papers and preprints during the coronavirus disease 2019 pandemic:
Progress or problems with peer review? Advances in Chronic Kidney
Disease. 2020;27: 418--426.
doi:\href{https://doi.org/10.1053/j.ackd.2020.08.003}{10.1053/j.ackd.2020.08.003}}

\leavevmode\hypertarget{ref-Anazco2021publication}{}%
\CSLLeftMargin{4. }
\CSLRightInline{Añazco D, Nicolalde B, Espinosa I, Camacho J, Mushtaq M,
Gimenez J, et al. Publication rate and citation counts for preprints
released during the COVID-19 pandemic: The good, the bad and the ugly.
PeerJ. 2021;9: e10927.
doi:\href{https://doi.org/10.7717/peerj.10927}{10.7717/peerj.10927}}

\leavevmode\hypertarget{ref-Bramstedt2020carnage}{}%
\CSLLeftMargin{5. }
\CSLRightInline{Bramstedt KA. The carnage of substandard research during
the COVID-19 pandemic: A call for quality. Journal of Medical Ethics.
2020;46: 803--807.
doi:\href{https://doi.org/10.1136/medethics-2020-106494}{10.1136/medethics-2020-106494}}

\leavevmode\hypertarget{ref-Broggini2017reproducible}{}%
\CSLLeftMargin{6. }
\CSLRightInline{Broggini F, Dellinger J, Fomel S, Liu Y. Reproducible
research: Geophysics papers of the future - introduction. Geophysics.
2017;82.
doi:\href{https://doi.org/10.1190/geo2017-0918-spseintro.1}{10.1190/geo2017-0918-spseintro.1}}

\leavevmode\hypertarget{ref-Ince2012case}{}%
\CSLLeftMargin{7. }
\CSLRightInline{Ince DC, Hatton L, Graham-Cumming J. The case for open
computer programs. Nature. 2012;482: 485--488.
doi:\href{https://doi.org/10.1038/nature10836}{10.1038/nature10836}}

\leavevmode\hypertarget{ref-Ioannidis2014increasing}{}%
\CSLLeftMargin{8. }
\CSLRightInline{Ioannidis JPA, Greenland S, Hlatky MA, Khoury MJ,
Macleod MR, Moher D, et al. Increasing value and reducing waste in
research design, conduct, and analysis. Lancet. 2014;383: 166--175.
doi:\href{https://doi.org/10.1016/s0140-6736(13)62227-8}{10.1016/s0140-6736(13)62227-8}}

\leavevmode\hypertarget{ref-Brunsdon2020opening}{}%
\CSLLeftMargin{9. }
\CSLRightInline{Brunsdon C, Comber A. Opening practice: Supporting
reproducibility and critical spatial data science. Journal of
Geographical Systems. 2020;
doi:\href{https://doi.org/10.1007/s10109-020-00334-2}{10.1007/s10109-020-00334-2}}

\leavevmode\hypertarget{ref-Konkol2019examination}{}%
\CSLLeftMargin{10. }
\CSLRightInline{Konkol M, Kray C. In-depth examination of spatiotemporal
figures in open reproducible research. Cartography and Geographic
Information Science. 2019;46: 412--427.
doi:\href{https://doi.org/10.1080/15230406.2018.1512421}{10.1080/15230406.2018.1512421}}

\leavevmode\hypertarget{ref-Konkol2019computational}{}%
\CSLLeftMargin{11. }
\CSLRightInline{Konkol M, Kray C, Pfeiffer M. Computational
reproducibility in geoscientific papers: Insights from a series of
studies with geoscientists and a reproduction study. International
Journal of Geographical Information Science. 2019;33: 408--429.
doi:\href{https://doi.org/10.1080/13658816.2018.1508687}{10.1080/13658816.2018.1508687}}

\leavevmode\hypertarget{ref-Iqbal2016reproducible}{}%
\CSLLeftMargin{12. }
\CSLRightInline{Iqbal SA, Wallach JD, Khoury MJ, Schully SD, Ioannidis
JPA. Reproducible research practices and transparency across the
biomedical literature. Plos Biology. 2016;14.
doi:\href{https://doi.org/10.1371/journal.pbio.1002333}{10.1371/journal.pbio.1002333}}

\leavevmode\hypertarget{ref-Stodden2018empirical}{}%
\CSLLeftMargin{13. }
\CSLRightInline{Stodden V, Seiler J, Ma ZK. An empirical analysis of
journal policy effectiveness for computational reproducibility.
Proceedings of the National Academy of Sciences of the United States of
America. 2018;115: 2584--2589.
doi:\href{https://doi.org/10.1073/pnas.1708290115}{10.1073/pnas.1708290115}}

\leavevmode\hypertarget{ref-Sumner2020reproducibility}{}%
\CSLLeftMargin{14. }
\CSLRightInline{Sumner J, Haynes L, Nathan S, Hudson-Vitale C, McIntosh
LD. Reproducibility and reporting practices in COVID-19 preprint
manuscripts. medRxiv. 2020; 2020.03.24.20042796.
doi:\href{https://doi.org/10.1101/2020.03.24.20042796}{10.1101/2020.03.24.20042796}}

\leavevmode\hypertarget{ref-Gustot2020quality}{}%
\CSLLeftMargin{15. }
\CSLRightInline{Gustot T. Quality and reproducibility during the
COVID-19 pandemic. JHEP Rep. 2020;2: 100141.
doi:\href{https://doi.org/10.1016/j.jhepr.2020.100141}{10.1016/j.jhepr.2020.100141}}

\leavevmode\hypertarget{ref-Herndon2014high}{}%
\CSLLeftMargin{16. }
\CSLRightInline{Herndon T, Ash M, Pollin R. Does high public debt
consistently stifle economic growth? A critique of reinhart and rogoff.
Cambridge Journal of Economics. 2014;38: 257--279.
doi:\href{https://doi.org/10.1093/cje/bet075}{10.1093/cje/bet075}}

\leavevmode\hypertarget{ref-Basu2017ten}{}%
\CSLLeftMargin{17. }
\CSLRightInline{Basu S, Carney MA, Kenworthy NJ. Ten years after the
financial crisis: The long reach of austerity and its global impacts on
health. Social Science \& Medicine. 2017;187: 203--207.
doi:\href{https://doi.org/10.1016/j.socscimed.2017.06.026}{10.1016/j.socscimed.2017.06.026}}

\leavevmode\hypertarget{ref-Florida2020how}{}%
\CSLLeftMargin{18. }
\CSLRightInline{Florida R, Glaeser E, Sharif M, Bedi K, Campanella T,
Chee C, et al. How life in our cities will look after the coronavirus
pandemic. Foreign Policy. 2020;1. Available:
\url{https://foreignpolicy.com/2020/05/01/future-of-cities-urban-life-after-coronavirus-pandemic/}}

\leavevmode\hypertarget{ref-Sy2021population}{}%
\CSLLeftMargin{19. }
\CSLRightInline{Sy KTL, White LF, Nichols BE. Population density and
basic reproductive number of COVID-19 across united states counties.
PLOS ONE. 2021;16: e0249271.
doi:\href{https://doi.org/10.1371/journal.pone.0249271}{10.1371/journal.pone.0249271}}

\leavevmode\hypertarget{ref-Moore1970urban}{}%
\CSLLeftMargin{20. }
\CSLRightInline{Moore EG, Brown LA. Urban acquaintance fields: An
evaluation of a spatial model. Environment and Planning. 1970;2:
443--454. Available:
\url{http://www.envplan.com/abstract.cgi?id=a020443}}

\leavevmode\hypertarget{ref-Moore1970some}{}%
\CSLLeftMargin{21. }
\CSLRightInline{Moore EG. Some spatial properties of urban contact
fields. Geographical Analysis. 1970;2: 376--386. }

\leavevmode\hypertarget{ref-Farber2011running}{}%
\CSLLeftMargin{22. }
\CSLRightInline{Farber S, Páez A. Running to stay in place: The time-use
implications of automobile oriented land-use and travel. Journal of
Transport Geography. 2011;19: 782--793.
doi:\href{https://doi.org/10.1016/j.jtrangeo.2010.09.008}{10.1016/j.jtrangeo.2010.09.008}}

\leavevmode\hypertarget{ref-Rocklov2020high}{}%
\CSLLeftMargin{23. }
\CSLRightInline{Rocklöv J, Sjödin H. High population densities catalyse
the spread of COVID-19. Journal of Travel Medicine. 2020;27.
doi:\href{https://doi.org/10.1093/jtm/taaa038}{10.1093/jtm/taaa038}}

\leavevmode\hypertarget{ref-Li2018effect}{}%
\CSLLeftMargin{24. }
\CSLRightInline{Li R, Richmond P, Roehner BM. Effect of population
density on epidemics. Physica A: Statistical Mechanics and its
Applications. 2018;510: 713--724.
doi:\href{https://doi.org/10.1016/j.physa.2018.07.025}{10.1016/j.physa.2018.07.025}}

\leavevmode\hypertarget{ref-Cruz2020exploring}{}%
\CSLLeftMargin{25. }
\CSLRightInline{Cruz CJP, Ganly R, Li Z, Gietel-Basten S. Exploring the
young demographic profile of COVID-19 cases in hong kong: Evidence from
migration and travel history data. PLOS ONE. 2020;15: e0235306.
doi:\href{https://doi.org/10.1371/journal.pone.0235306}{10.1371/journal.pone.0235306}}

\leavevmode\hypertarget{ref-Micallef2020first}{}%
\CSLLeftMargin{26. }
\CSLRightInline{Micallef S, Piscopo TV, Casha R, Borg D, Vella C, Zammit
M-A, et al. The first wave of COVID-19 in malta; a national
cross-sectional study. PLOS ONE. 2020;15: e0239389.
doi:\href{https://doi.org/10.1371/journal.pone.0239389}{10.1371/journal.pone.0239389}}

\leavevmode\hypertarget{ref-Gomez2021infekta}{}%
\CSLLeftMargin{27. }
\CSLRightInline{Gomez J, Prieto J, Leon E, Rodríguez A. INFEKTA---an
agent-based model for transmission of infectious diseases: The COVID-19
case in bogotá, colombia. PLOS ONE. 2021;16: e0245787.
doi:\href{https://doi.org/10.1371/journal.pone.0245787}{10.1371/journal.pone.0245787}}

\leavevmode\hypertarget{ref-Noland1995perceived}{}%
\CSLLeftMargin{28. }
\CSLRightInline{Noland RB. PERCEIVED RISK AND MODAL CHOICE - RISK
COMPENSATION IN TRANSPORTATION SYSTEM. Accident Analysis and Prevention.
1995;27: 503--521.
doi:\href{https://doi.org/10.1016/0001-4575(94)00087-3}{10.1016/0001-4575(94)00087-3}}

\leavevmode\hypertarget{ref-Richens2000condoms}{}%
\CSLLeftMargin{29. }
\CSLRightInline{Richens J, Imrie J, Copas A. Condoms and seat belts: The
parallels and the lessons. Lancet. 2000;355: 400--403.
doi:\href{https://doi.org/10.1016/s0140-6736(99)09109-6}{10.1016/s0140-6736(99)09109-6}}

\leavevmode\hypertarget{ref-Phillips2011risk}{}%
\CSLLeftMargin{30. }
\CSLRightInline{Phillips RO, Fyhri A, Sagberg F. Risk compensation and
bicycle helmets. Risk Analysis. 2011;31: 1187--1195.
doi:\href{https://doi.org/10.1111/j.1539-6924.2011.01589.x}{10.1111/j.1539-6924.2011.01589.x}}

\leavevmode\hypertarget{ref-Chauhan2021covid}{}%
\CSLLeftMargin{31. }
\CSLRightInline{Chauhan RS, Capasso da Silva D, Salon D, Shamshiripour
A, Rahimi E, Sutradhar U, et al. COVID-19 related attitudes and risk
perceptions across urban, rural, and suburban areas in the united
states. Findings. Network Design Lab; 2021;
doi:\href{https://doi.org/10.32866/001c.23714}{10.32866/001c.23714}}

\leavevmode\hypertarget{ref-Paez2020spatio}{}%
\CSLLeftMargin{32. }
\CSLRightInline{Paez A, Lopez FA, Menezes T, Cavalcanti R, Pitta MG da
R. A spatio-temporal analysis of the environmental correlates of
COVID-19 incidence in spain. Geographical Analysis. 2020;n/a.
doi:\href{https://doi.org/10.1111/gean.12241}{10.1111/gean.12241}}

\leavevmode\hypertarget{ref-Hamidi2020density}{}%
\CSLLeftMargin{33. }
\CSLRightInline{Hamidi S, Sabouri S, Ewing R. Does density aggravate the
COVID-19 pandemic? Journal of the American Planning Association.
2020;86: 495--509.
doi:\href{https://doi.org/10.1080/01944363.2020.1777891}{10.1080/01944363.2020.1777891}}

\leavevmode\hypertarget{ref-Jamal2020Changes}{}%
\CSLLeftMargin{34. }
\CSLRightInline{Jamal S, Paez A. Changes in trip-making frequency by
mode during COVID-19. Findings. Network Design Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.17977}{10.32866/001c.17977}}

\leavevmode\hypertarget{ref-Harris2021Changes}{}%
\CSLLeftMargin{35. }
\CSLRightInline{Harris MA, Branion-Calles M. Changes in commute mode
attributed to COVID-19 risk in canadian national survey data. Findings.
Network Design Lab; 2021;
doi:\href{https://doi.org/10.32866/001c.19088}{10.32866/001c.19088}}

\leavevmode\hypertarget{ref-Molloy2020Tracing}{}%
\CSLLeftMargin{36. }
\CSLRightInline{Molloy J, Tchervenkov C, Hintermann B, Axhausen KW.
Tracing the sars-CoV-2 impact: The first month in switzerland. Findings.
Network Design Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.12903}{10.32866/001c.12903}}

\leavevmode\hypertarget{ref-Feyman2020effectiveness}{}%
\CSLLeftMargin{37. }
\CSLRightInline{Feyman Y, Bor J, Raifman J, Griffith KN. Effectiveness
of COVID-19 shelter-in-place orders varied by state. PLOS ONE. 2020;15:
e0245008.
doi:\href{https://doi.org/10.1371/journal.pone.0245008}{10.1371/journal.pone.0245008}}

\leavevmode\hypertarget{ref-Praharaj2020Using}{}%
\CSLLeftMargin{38. }
\CSLRightInline{Praharaj S, King D, Pettit C, Wentz E. Using aggregated
mobility data to measure the effect of COVID-19 policies on mobility
changes in sydney, london, phoenix, and pune. Findings. Network Design
Lab; 2020;
doi:\href{https://doi.org/10.32866/001c.17590}{10.32866/001c.17590}}

\leavevmode\hypertarget{ref-Paez2020using}{}%
\CSLLeftMargin{39. }
\CSLRightInline{Paez A. Using google community mobility reports to
investigate the incidence of COVID-19 in the united states. Findings.
2020; doi:\url{https://doi.org/10.32866/001c.12976}}

\leavevmode\hypertarget{ref-Noland2021mobility}{}%
\CSLLeftMargin{40. }
\CSLRightInline{Noland RB. Mobility and the effective reproduction rate
of COVID-19. Journal of Transport \& Health. 2021;20: 101016.
doi:\url{https://doi.org/10.1016/j.jth.2021.101016}}

\leavevmode\hypertarget{ref-Bhadra2021impact}{}%
\CSLLeftMargin{41. }
\CSLRightInline{Bhadra A, Mukherjee A, Sarkar K. Impact of population
density on covid-19 infected and mortality rate in india. Modeling Earth
Systems and Environment. 2021;7: 623--629.
doi:\href{https://doi.org/10.1007/s40808-020-00984-7}{10.1007/s40808-020-00984-7}}

\leavevmode\hypertarget{ref-Kadi2020population}{}%
\CSLLeftMargin{42. }
\CSLRightInline{Kadi N, Khelfaoui M. Population density, a factor in the
spread of COVID-19 in algeria: Statistic study. Bulletin of the National
Research Centre. 2020;44.
doi:\href{https://doi.org/10.1186/s42269-020-00393-x}{10.1186/s42269-020-00393-x}}

\leavevmode\hypertarget{ref-Pequeno2020air}{}%
\CSLLeftMargin{43. }
\CSLRightInline{Pequeno P, Mendel B, Rosa C, Bosholn M, Souza JL,
Baccaro F, et al. Air transportation, population density and temperature
predict the spread of COVID-19 in brazil. PeerJ. 2020;8: e9322.
doi:\href{https://doi.org/10.7717/peerj.9322}{10.7717/peerj.9322}}

\leavevmode\hypertarget{ref-Fielding2020social}{}%
\CSLLeftMargin{44. }
\CSLRightInline{Fielding-Miller RK, Sundaram ME, Brouwer K. Social
determinants of COVID-19 mortality at the county level. PLOS ONE.
2020;15: e0240151.
doi:\href{https://doi.org/10.1371/journal.pone.0240151}{10.1371/journal.pone.0240151}}

\leavevmode\hypertarget{ref-Roy2020factors}{}%
\CSLLeftMargin{45. }
\CSLRightInline{Roy S, Ghosh P. Factors affecting COVID-19 infected and
death rates inform lockdown-related policymaking. PLOS ONE. 2020;15:
e0241165.
doi:\href{https://doi.org/10.1371/journal.pone.0241165}{10.1371/journal.pone.0241165}}

\leavevmode\hypertarget{ref-Wong2020spreading}{}%
\CSLLeftMargin{46. }
\CSLRightInline{Wong DWS, Li Y. Spreading of COVID-19: Density matters.
PLOS ONE. 2020;15: e0242398.
doi:\href{https://doi.org/10.1371/journal.pone.0242398}{10.1371/journal.pone.0242398}}

\leavevmode\hypertarget{ref-Sun2020impacts}{}%
\CSLLeftMargin{47. }
\CSLRightInline{Sun Z, Zhang H, Yang Y, Wan H, Wang Y. Impacts of
geographic factors and population density on the COVID-19 spreading
under the lockdown policies of china. Science of The Total Environment.
2020;746: 141347.
doi:\href{https://doi.org/10.1016/j.scitotenv.2020.141347}{10.1016/j.scitotenv.2020.141347}}

\leavevmode\hypertarget{ref-Skorka2020macroecology}{}%
\CSLLeftMargin{48. }
\CSLRightInline{Skórka P, Grzywacz B, Moroń D, Lenda M. The macroecology
of the COVID-19 pandemic in the anthropocene. PLOS ONE. 2020;15:
e0236856.
doi:\href{https://doi.org/10.1371/journal.pone.0236856}{10.1371/journal.pone.0236856}}

\leavevmode\hypertarget{ref-White2020state}{}%
\CSLLeftMargin{49. }
\CSLRightInline{White ER, Hébert-Dufresne L. State-level variation of
initial COVID-19 dynamics in the united states. PLOS ONE. 2020;15:
e0240648.
doi:\href{https://doi.org/10.1371/journal.pone.0240648}{10.1371/journal.pone.0240648}}

\leavevmode\hypertarget{ref-Khavarian2021high}{}%
\CSLLeftMargin{50. }
\CSLRightInline{Khavarian-Garmsir AR, Sharifi A, Moradpour N. Are
high-density districts more vulnerable to the COVID-19 pandemic?
Sustainable Cities and Society. 2021;70: 102911.
doi:\href{https://doi.org/10.1016/j.scs.2021.102911}{10.1016/j.scs.2021.102911}}

\leavevmode\hypertarget{ref-Hamidi2020longitudinal}{}%
\CSLLeftMargin{51. }
\CSLRightInline{Hamidi S, Ewing R, Sabouri S. Longitudinal analyses of
the relationship between development density and the COVID-19 morbidity
and mortality rates: Early evidence from 1,165 metropolitan counties in
the united states. Health \& Place. 2020;64: 102378.
doi:\href{https://doi.org/10.1016/j.healthplace.2020.102378}{10.1016/j.healthplace.2020.102378}}

\leavevmode\hypertarget{ref-Lee2020human}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Lee M, Zhao J, Sun Q, Pan Y, Zhou W, Xiong C, et al.
Human mobility trends during the early stage of the COVID-19 pandemic in
the united states. PLOS ONE. 2020;15: e0241468.
doi:\href{https://doi.org/10.1371/journal.pone.0241468}{10.1371/journal.pone.0241468}}

\leavevmode\hypertarget{ref-Amadu2021assessing}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Amadu I, Ahinkorah BO, Afitiri A-R, Seidu A-A, Ameyaw
EK, Hagan JE, et al. Assessing sub-regional-specific strengths of
healthcare systems associated with COVID-19 prevalence, deaths and
recoveries in africa. PLOS ONE. 2021;16: e0247274.
doi:\href{https://doi.org/10.1371/journal.pone.0247274}{10.1371/journal.pone.0247274}}

\leavevmode\hypertarget{ref-Feng2020spread}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Feng Y, Li Q, Tong X, Wang R, Zhai S, Gao C, et al.
Spatiotemporal spread pattern of the COVID-19 cases in china. PLOS ONE.
2020;15: e0244351.
doi:\href{https://doi.org/10.1371/journal.pone.0244351}{10.1371/journal.pone.0244351}}

\leavevmode\hypertarget{ref-Inbaraj2021seroprevalence}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Inbaraj LR, George CE, Chandrasingh S. Seroprevalence of
COVID-19 infection in a rural district of south india: A
population-based seroepidemiological study. PLOS ONE. 2021;16: e0249247.
doi:\href{https://doi.org/10.1371/journal.pone.0249247}{10.1371/journal.pone.0249247}}

\leavevmode\hypertarget{ref-Souris2020covid}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Souris M, Gonzalez J-P. COVID-19: Spatial analysis of
hospital case-fatality rate in france. PLOS ONE. 2020;15: e0243606.
doi:\href{https://doi.org/10.1371/journal.pone.0243606}{10.1371/journal.pone.0243606}}

\leavevmode\hypertarget{ref-Ahmad2020association}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Ahmad K, Erqou S, Shah N, Nazir U, Morrison AR,
Choudhary G, et al. Association of poor housing conditions with COVID-19
incidence and mortality across US counties. PLOS ONE. 2020;15: e0241327.
doi:\href{https://doi.org/10.1371/journal.pone.0241327}{10.1371/journal.pone.0241327}}

\leavevmode\hypertarget{ref-Noury2021how}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Noury A, François A, Gergaud O, Garel A. How does
COVID-19 affect electoral participation? Evidence from the french
municipal elections. PLOS ONE. 2021;16: e0247026.
doi:\href{https://doi.org/10.1371/journal.pone.0247026}{10.1371/journal.pone.0247026}}

\leavevmode\hypertarget{ref-Wang2021transmission}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Wang F, Tan Z, Yu Z, Yao S, Guo C. Transmission and
control pressure analysis of the COVID-19 epidemic situation using
multisource spatio-temporal big data. PLOS ONE. 2021;16: e0249145.
doi:\href{https://doi.org/10.1371/journal.pone.0249145}{10.1371/journal.pone.0249145}}

\leavevmode\hypertarget{ref-Brandtner2021creatures}{}%
\CSLLeftMargin{60. }
\CSLRightInline{Brandtner C, Bettencourt LMA, Berman MG, Stier AJ.
Creatures of the state? Metropolitan counties compensated for state
inaction in initial u.s. Response to COVID-19 pandemic. PLOS ONE.
2021;16: e0246249.
doi:\href{https://doi.org/10.1371/journal.pone.0246249}{10.1371/journal.pone.0246249}}

\leavevmode\hypertarget{ref-Stephens2021impact}{}%
\CSLLeftMargin{61. }
\CSLRightInline{Stephens KE, Chernyavskiy P, Bruns DR. Impact of
altitude on COVID-19 infection and death in the united states: A
modeling and observational study. PLOS ONE. 2021;16: e0245055.
doi:\href{https://doi.org/10.1371/journal.pone.0245055}{10.1371/journal.pone.0245055}}

\leavevmode\hypertarget{ref-Maddala1983limited}{}%
\CSLLeftMargin{62. }
\CSLRightInline{Maddala GS. Limited-dependent and qualitative variables
in econometrics. Cambridge: Cambridge University Press; 1983. }

\leavevmode\hypertarget{ref-Cressie1989geostatistics}{}%
\CSLLeftMargin{63. }
\CSLRightInline{Cressie N. Geostatistics. The American Statistician.
1989;43: 197.
doi:\href{https://doi.org/10.2307/2685361}{10.2307/2685361}}

\end{CSLReferences}

\nolinenumbers

\end{document}