### 2021-05-25 R0-Density-Reanalysis/R0-Density-Reanalysis-Title-Page.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Science of the Total Environment} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The importance of reproducibility in COVID-19 research: the case of population density and the spread of the pandemic},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}

% Pandoc citation processing

% Pandoc header
\usepackage{setspace}\doublespacing

\begin{document}
\begin{frontmatter}

  \title{The importance of reproducibility in COVID-19 research: the
case of population density and the spread of the pandemic}
    \author[McMaster University]{Antonio Paez\corref{1}}
   \ead{paezha@mcmaster.ca} 
      \address[McMaster University]{School of Earth, Environment and
Society, 1280 Main St West, Hamilton, Ontario L8S 4K1 Canada}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  The emergence of the novel SARS-CoV-2 coronavirus and the global
  COVID-19 pandemic has led to explosive growth in scientific research.
  Given the high stakes of the situation, it is essential that
  scientific activities, on which good policy depends, are as
  transparent and reproducible as possible. Reproducibility is key for
  the efficient operation of the self-correction mechanisms of science,
  which work to weed out errors and refine our understanding of social
  and physical phenomena. In this paper, the importance of
  reproducibility is illustrated for the case of the association between
  population density and the the spread of SARS-CoV-2. Transparency and
  openness means that the same problem can, with relatively modest
  efforts, be examined by independent researchers who can verify
  findings, and bring to bear different perspectives, approaches, and
  methods---sometimes with consequential changes in the conclusions, as
  the empirical example in this paper shows.
  \end{abstract}

 \end{frontmatter}

\end{document}