### 2021-05-26 R0-Density-Reanalysis/Highlights.Rmd

---
title: "Highlights"
output:
  pdf_document: default
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

- The emerging literature on the relationship between population density and the spread of COVID-19 is reviewed.

- It is shown that reproducibility in this literature remains an elusive goal.

- Analysis of population density and basic reproductive number shows that density is not destiny.

- The reanalysis and reappraisal of the conclusions demonstrates the importance of reproducible research.

- The paper is an example of a reproducible research project